---
title: "Correlation between Trio and Shank3"
output: html_notebook
---


```{r}
trio.xls <- read_xlsx("../data/results_PSD-Trio-proteome_Two-sample_mod_T_2020-07-24.xlsx")
trio.xls <- trio.xls %>% drop_na(logFC.WT.over.KO)
shank3.xls <- read_xlsx("../data/results_PSD-Shank3-proteome_Two-sample_mod_T_2020-07-22.xlsx")
shank3.xls <- shank3.xls %>% drop_na(logFC.WT.over.KO)
```
The number of proteins measured in Trio PSD is:
```{r}
dim(trio.xls)[1]
```
The number of proteins measured in Shank3 PSD is:
```{r}
dim(shank3.xls)[1]
```

The number of proteins measured across both experiments is:
```{r}
sharedIds <- intersect(trio.xls$id, shank3.xls$id)
length(sharedIds)
```
699 proteins measured in Trio PSD but not Shank3. 1534 proteins measured in Shank3 PSD but not Trio.
```{r}
length(setdiff(trio.xls$id, shank3.xls$id))
length(setdiff(shank3.xls$id, trio.xls$id))
```
```{r}
#trio.shared <- trio.xls %>% subset(id %in% sharedIds)
#shank3.shared <- shank3.xls %>% subset(id %in% sharedIds)
trio.xls <- trio.xls %>% dplyr::mutate(logFC.KO.over.WT = logFC.WT.over.KO*(-1))
shank3.xls <- shank3.xls %>% dplyr::mutate(logFC.KO.over.WT = logFC.WT.over.KO*(-1))
joint <- inner_join(trio.xls, shank3.xls, by='id')
lim <- c(-2,2)
plot(joint$logFC.KO.over.WT.x, joint$logFC.KO.over.WT.y, xlim=lim, ylim=lim)
cor(joint$logFC.KO.over.WT.x, joint$logFC.KO.over.WT.y)
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

